---
title: "Diamond Assignment"
output: html_document
author: "Yousuf Ahmed Nahian" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We want to study the `diamonds` dataset.

```{r}
library(ggplot2)
library(dplyr)
```

#### 1. Display the first few rows of the dataset.

```{r}
head(diamonds)
```

#### 2. The `cut` variable

```{r}
levels(diamonds$cut)

table(diamonds$cut)
```

2.1 Determine the type of the variable `cut` using `class()`.

```{r}
class(diamonds$cut)
```

2.2 What happens if you use `typeof()` instead? Do you have an idea why?

```{r}
typeof(diamonds$cut)
```

class() = What kind of data is this for analysis?
typeof() = How is it stored in memory?

#### 3. In which package is the dataset diamonds contained in? How can you look it up?

Diamonds Dataset is contained at ggplot2 package.
Using ´help("diamonds") we can look up which package it belongs to.

#### 4. Create a scatterplot to study the relationship between `carat` and `price`. Add a linear trendline. Is the relationship positive or negative?

```{r}
library(ggplot2)
ggplot(diamonds, aes (x = carat, y = price)) + geom_smooth(method = "lm", col = "blue") + labs(title = "Relationship Between Carat And Price", x = "Carat", y = "Price")
```
From the graph we can see that as carat increases the price also increases. So, the relationship is Positive.


#### 5. Create a boxplot to visualize the distribution of price across different cuts. Are there patterns that seem surprising? Suggest possible explanations.

```{r}
library(ggplot2)
ggplot(diamonds, aes(x = cut, y = price)) + geom_boxplot(fill = "yellow") + labs(title = "The distribution of price across different cuts", x = "Cut", y= "Price")
```
So patterns seems surprising.
Possible Explanations -

1.Price Depends on multiple factors, not just on cut.

2.Some lower carat diamonds can be more expensive due to its carat, rarity or clarity.

#### 6. Create scatterplots to investigate how the relationship between carat and price differs depending on the cut of the diamond. Write \~3 sentences to describe similarities and differences you observe between cuts.
```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Carat vs Price by Cut",
       x = "Carat",
       y = "Price")
```
All cuts show positive relationship.

Premium and Ideal tend to have higher prices for the same carat compared to other.

The spread of prices is wider for lower cuts, suggesting more variability in price for fair or good diamonds.

#### True or False?

Are the following statements about the `diamonds` dataset true or false? Show code to justify your answer.

7.  The diamonds dataset contains information on about 5400 diamonds.

False
```{r}
dim(diamonds)
```

8.  Most diamonds in the dataset are priced below \$5000.

True
```{r}
sum(diamonds$price < 5000)
```

9.  The `geom_histogram()` function is used to show relationships between two numeric variables. (no code needed)

False